Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.

### 2024-03-01 docs/_tex/index.tex
Pierre comments

Coastal areas are vital hotshots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. These meadows offer a
myriad of ecosystem services to humanity, including limitation of
coastline erosion, reducing the risk of eutrophication, carbon
sequestration, and oxygen production. They serve as vital habitats for a
diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to their proximity to human activities, seagrass meadows are directly
exposed to and impacted by anthropogenic pressures. Global regression
and fragmentation are currently observed due to diseases, disasters,
coastal urbanization, sea reclamation, as well as fishing activities,
dredging, sea level rise, coastal erosion, competition with alien
species, and reduction in water quality \citetext{\citealp[
]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in European
sites, allowing an overall recovery of seagrass ecosystems at the local
scale, many other coastal waters worldwide are still subjected to strong
eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp{Zoffoli2021}}. Coastal eutrophication has been
associated to anomalous accumulation of green macroalgae, the so-called
green tides. Green tides produce shade and suffication over seagrass
individuals, thus threatening the health of seagrass ecosystems
\citetext{\citealp[ ]{Duarte2002}; \citealp{wang2022}}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of global and regional
monitoring programs for systematically surveying different Essential
Oceanic Variable \citep{Miloslavich2018} as seagrass coverage and
composition; as well as Essential Biodiversity Variable
\citep{Pereira2013} such as seagrass taxonomic diversity; species
distribution, population abundance, and phenology. Monitoring programs
also prioritize the identification of threats to these ecosystems,
particularly during early stages, to facilitate effective mitigation
actions. Traditionally, these ecological parameters have been quantified
through in situ measurements, although this approach faces several
constraints over intertidal zones. Intertidal meadows are only partially
exposed during low tide and can be situated in difficult-to-reach
mudflats, potentially leading to inaccurate and limited estimations with
conventional sampling techniques \citep{nijland2019}. However, satellite
data have been proven effective in complenmenting in situ surveys,
allowing for the rapid and consitent retrieval of EOV's over extensive
seagrass meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

Satellite remote sensing offers the advantage of acquiring large-scale
data in real-time but presents its inherent challenges. Free access
satellite data (e.g., Sentinel-2 and Landsat8/9) provide relatively low
spatial resolution data (10 - 30 m) across a limited number of spectral
bands. These characteristics can be a limitation to accurately
discriminating seagrass from others co-existing macrophytes over the
meadow. Chlorophyceae (Green Algae) and marine Magnoliopsida (Seagrass)
share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. As a result, their respective
spectral signatures can be considered similar by a non-expert observer
\citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}. Recently,
using a hyperspectral library, \citep{Davies2023} demonstrated that the
spectral resolution of Sentinel-2, might be enough for the
discrimination between Magnoliopsida and Chlorophyceae. However, green
tide events occur at small spatial scales that are not observable using
satellite imagery \citep{tuya2013}, especially during the initial stage
of the event.

Remote sensing drone acquisitions are presented as a tool that can
potentially fill gaps left by satellite and in situ data. Drone can
cover large expanses while recording imagery at significantly higher
spatial resolutions than satellite (pixel size from cm to mm) and still
capturing data at multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping the spatial
distribution of species \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, many studies showcase their findings with
study case limited to a single flight, restricting the generalizability
of their application to other sites \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. This study aimed
to analyze the potential of a drone equipped with a multispectral sensor
for mapping intertidal macrophytes, with a particular focus on
discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were
performed over soft-bottom intertidal areas along two European countries
(France and Portugal), covering a wide range of habitats, from
monospecific seagrass meadows to meadows mixed with green or red algae.
A deep learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without a
loss of accuracy in predictions.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connect to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : One is located in one of these
island (Arz island). The other one is located in the sourthern part of
the Gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. With up to 6m in tidal range
during spring tide, a 6 km² seagrass meadow can uncover at low tide. The
sites we surveyed (L'Epine and Barbatre, Figure~\ref{fig-map} B) are
located in the southernmost part of the bay, along the Noirmoutier
Island, and host a monospecific \emph{Zostera noltei} seagrass bed with
very little mixing with other macrophytes. Three sites have been survey
in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The
extent of this lagoon is about 80 km² with many narrow channels, large
salt marshes and many mudflats that uncovers at low tide. It is
connected to the open sea through a single channel, with a tidal lag
between the North and the South of the lagoon. The southernmost site
(Gafanha) is a mudflat located in an affluent of the lagoon whereas the
two other sites are situated in the middle of the lagoon and are only
accessible by boat (Figure~\ref{fig-map} C). These Portuguese sites were
characterized by a more diverse intertidal area, where the seagrass
meadow could intermingle with red algae (Rhodophyceae), brown algae
(Phaeophyceae), or green algae (Chlorophyceae).

@@

\caption{\label{fig-map}Sites where the different drone flights were
made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon.
Light green represent terrestrial areas whereas darkgreen represent
intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from blue
(444 nm) to near infrared (840 nm). An angle of 90° has been maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image has been set. Images have been
calibrated in reflectance using a calibration panel reflective at
\textasciitilde50\% provided by the camera's manufaturer and a
Downwelling Light Sensor (DLS2) that was utilized to acquire irradiance
data during the flight. A Structure-from-Motion photogrammetry software
(Agisoft Metashape) has been used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking has remained consistent for every flight. First, tying
key points were detected inside of each image and between overlapping
images in order to obtain a sparse point cloud. This cloud was cleaned
using reprojection accuracy metric in order to remove noisy points. A
dense point cloud has been then produced using a structure from motion
algorithm. A surface interpolation of this dense point cloud has been
made to obtain a Digital Surface Model (DSM), used to reconstruct the
multispectral ortho-image. Across all the sites, flights were made at
two different altitudes : 12 m or/and 120 m. Low altitude drone flights
produce ortho-images with a very high spatial resolution (8 mm per
pixel), making it simple to visually distinguish between the various
types of vegetation. High altitude fights on the other hand allow to
cover large areas and produced images with a pixel size of 80 mm (Table
1).

@@

\begin{longtable}[c]{ccccc}
\caption{List of drone Flight, summarising the date, the altitude and the purpose
of each flight.}\tabularnewline

@@

Before each flight, targets used as Ground Control Points (GCPs) were
dispersed and georeferenced with a Trimble © Geo XH 6000 differential
GPS (dGPS). GCPs were used to correct georeferencing imprecision of
orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was
also used to georeference quardrats of 0.25cm² used to assess the
presence or absence of 5 key classes of vegetation : Bacillariophyceae
(Diatoms forming large biofilm over the sediment during low tide),
Phaeophyceae (Brown algae), Magnoliopsida (Seagrass), Chlorophyceae
(Green algae) and Rhodophyceae (Red algae)
(Figure~\ref{fig-vegetation}). Picture of each quadrat were taken and
uploaded on the online on the Global Biodiversity Information Facility
(GBIF) platform \citep{BedeGbif}. Each photograph was also processed to
estimate the percent cover of each type of vegetation. Georeferenced
hyperspectral reflectance signatures of each vegetation class were also
sampled using an ASD FieldSpec HandHeld 2 spectroradiomete, which
acquires reflectance between 325 and 1075 nm, with 1 nm of spectral
resolution. The methodological scheme followed in this study is
presented in Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-vegetation}Class of vegetation used to train the CNN
model and their standardised spectral signature}

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights were utilized for training and
prediction of the CNN model, whereas low-resolution flights were solely
employed for prediction purposes.}

@@

The spectral similarities of reflectance signatures between intertidal
green macrophytes (Magnoliopsida and Chlorophyceae) make their
discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was developed, trained, validated, and
applied to each drone flight (Figure~\ref{fig-workflow}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Convolutional Neural Network (CNN) model with 2 hidden
layers. The training pixels were categorized into 7 different classes,
representing the various habitats encountered at the different study
sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae,
Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were
used for training purposes because of their 8 mm spatial resolution
allowing to avoid spectral sub-pixel mixing and to accurately visually
differentiate various vegetation classes. More than 418,000 pixels at 8
mm resolution from the 3 training flights have been used to train the
model (Table 2). Twenty one variables were used by the model as
predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX
multispectral camera (ranging from 444 nm to 840 nm), the same 10
spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the NDVI
(Equation~\ref{eq-ndvi}). Standardisation of spectral bands is used to
eliminate the scaling differences between spectra and to limit the
effect of biomass on the shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

The CNN model was applied to all the 9 flights at both 12 and 120 m of
altitude. In situ information on georeferenced class type and percent
cover collected during each flight has been used to assess the CNN
accuracy. These images were used to construct a validation dataset
indicating the presence or absence of each class. Additionally to the
quadrat-based validation dataset, polygons of each class were photo
interpreted in order to increase the number of pixels of the validation
dataset. A confusion matrix, along with precision metrics such as the
global accuracy, the sensitivity, the specificity, and the Kappa
coefficient, was generated for each of the study sites. Finally each
validation matrix has been merge to create a unique matrix of the model.
A total of 536,000 pixels has been used to globally validate the CNN
model.

@@

A total of 9 prediction maps corresponding to the 9 drone flights were
obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanaha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the 5
vegetation classes on which the CNN model was trained, 4 were present on
this site. On this site, there is a mixture of Chlorophyceae and
Rhodophyceae over the seagrass meadow. This is also where
Bacillariophyceae is most abundant. Although the seagrass bed is solely
composed of \emph{Zostera noltei}, various colors can be observed: dark
green (indicating healthy beds with 100\% coverage) and whitish/brown
(indicating beds where leaves are dying). Regardless of the color, the
meadow is predicted as Magnoliopsida by the CNN model.

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
present in the flight, but it has been masked in the prediction map due
to glinting and misclassification caused by the turbid shallow water.
Most of the intertidal area has been classified as Magnoliopsida by the
model, even though there is discoloration of seagrass blades. Only a few
pixels have been classified as Chlorophyceae at this scale. Furthermore,
the area that was classified as Bacillariophyceae in the low-altitude
flight remains mostly classified as such in the high-altitude flight,
but some pixels of it are now classified as Magnoliopsida. The patches
of Rhodophyceae are still well classified. \emph{Spartina sp.} was
present in the northern part of the flight, near the land, and has been
classified either as Magnoliopsida or as Phaeophyceae.

@@

The High altitude flight made in the inner lagoon of the Ria de Aveiro
is the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). On this site, only seagrass and red algae where
seen on the field. The classification follows the same pattern, with a
patchy Magnoliopsida meadow mixed on the eastern part of the flight with
Rhodophyceae. As shown in the zoom (Figure~\ref{fig-Boat}), the edges of
the meadow can be colonised by Chlorophyceae.

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the flight from west to east. Alongside this dike, brown algae
attached to rocks and stranded green algae could be found. Despite the
high mixture between Chlorophyceae and Magnoliopsida in this flight,
these two classes are well discriminated by the classifier.

@@

A total amount of 536,000 pixels have been used to validate the Neural
Network classifier. The site with the least amount of validation data is
Kerdrean with a total of 5557 pixels whereas Marinha Lanzarote is the
site with the most amount of validation data with 159713 pixels. The
global accuracy of the model is 94.26\% and a Kappa coefficient of 0.92
(Figure~\ref{fig-Validation}). The least performing site is Gafanha High
with an accuracy of 75.45\% whereas Mataducos is the site with most
accurate prediction, with a global accuracy of 98.05\%. Overall
Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae are well
classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae is the least performing class with an
accuracy of 0.72 mainly due to a confusion with Magnoliopsida and
Sediment.

@@

The computation of the importance variable of each class indicate which
wavelength is mandatory to accurately predict the class
(Figure~\ref{fig-VIP}). Bands at 444 nm, 717 nm and 842 nm of the
Micasense camera are important for none of the vegetation classes. The
band at 531 nm is the only important predictor for the classifier to
accuratly predict chlorophyceae. In fact, at this wavelength, the
Chlorophyceae spectra has the maximum reflectance value of all the
classes. 531 nm and 740 nm are the most important predictors to predict
Phaeophyceae, corresponding to the minimum reflectance value among all
the classes. Bands at 475 and 560 nm are the most important predictors
for Bacillariophyceae and Rhodophyceae, respectively. 4 predictors,
ranging from the Green (560 nm) to the RedEdge (705 nm) are important to
predict accurately magnoliopsida.

@@

Figure~\ref{fig-upscaling} is showing the percent cover of a class
required to classify correctly the high altitude flight. When the
percent cover of the class is 100 \%, big pixels are well classified for
all the classes excepted for Bare Sediment, where it's well classified
80\% of the time. A vegetative percent cover of at least 80\% is need to
have all the big pixels well classified, at the exception of
Magnoliopsida that needs an higher percent cover (\textgreater90 \%) to
be well classified. Concerning the probability of each class, a really
high Percent cover is needed to confidently predict Bacillariophyceae.
To predict Chlorophyceae with a confidence of 0.85, a percent cover of
93 \% is needed, 90 \% for magnoliopsida, 92 \% for Rhodophyceae and 97
\% for Bacillariophyceae.

@@

The VIP analysis of the Neural Network model (see Fig. VIP) shows that
the 531 nm band is the most important spectral band for accurately
identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae
exhibits the highest reflectance among all other classes, highlighting
the difference in carotenoid proportion between seagrasses and green
algae.

>>>>

Coastal areas are vital hotshots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. Seagrass meadows provide a
myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production,protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been associated to anomalous accumulation of
green macroalgae, the so-called green tides (\citep{devlin2023nutrients}
; \citep{devlin2023nutrients}). Green tides produce shade and
suffication over seagrass individuals, thus threatening the health of
seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, populuation abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using in situ measurements. The acquisition
of field data in intertidal zones is however notoriously challenging.
Intertidal seagrass meadows are only partially exposed during low tide
and can be situated in difficult-to-reach mudflats, potentially leading
to inaccurate and limited estimations with conventional sampling
techniques \citep{nijland2019}. Satellite observations have been proven
effective in complementing in situ sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance measurements, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drone can potentially fill the gaps left by satellite remote sensing and
in situ measurments, due to their ability to provide spatially-explicit
observations at very high spatial resolutions (pixel size from cm to mm)
and still capturing data at multi-spectral resolution
\citetext{\citealp[ ]{fairley2022drone}; \citealp{oh2017use}}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management \citetext{\citealp[
]{adade2021}; \citealp[ ]{casella2020}; \citealp{angnuureng2022}} to
mapping species distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most sutdy cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green or red algae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connect to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the sourthern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red algae (Rhodophyceae), brown algae (Phaeophyceae),
and green algae (Chlorophyceae).

@@

\caption{\label{fig-map}Sites observed by drone. A: Gulf of Morbihan, B:
Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent
terrestrial areas whereas darkgreen represent intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each imagewas set for each flight. A downwelling
light sensor (DLS2) was used to acquire irradiance data concomitantly
with the camera measurements. Raw data were calibrated in reflectance
using a calibration panel reflective at \textasciitilde50\% provided by
the manufacturer. A structure-from-motion photogrammetry software
(Agisoft Metashape) was used to process images to obtain multispectral
orthomosaics of each flight. The workflow for orthomosaicking was the
same for every flight. First, tying key points were detected inside of
each image and between overlapping images in order to obtain a sparse
point cloud. This cloud was cleaned using reprojection accuracy metric
in order to remove noisy points. A dense point cloud was then produced
using a structure from motion algorithm. A surface interpolation of this
dense point cloud was made to obtain a digital surface model (DSM), used
to reconstruct the multispectral ortho-image. Across all sites, flights
were made at two different altitudes : 12 m or/and 120 m. Low altitude
drone flights produce ortho-images with a very high spatial resolution
(8 mm per pixel), making it simple to visually distinguish between the
various types of vegetation. High altitude fights on the other hand
allow to cover large areas and produced images with a pixel size of 80
mm (Table 1).

@@

\begin{longtable}[c]{ccccc}
\caption{List of drone flight, summarising the date, the altitude and the purpose
of each flight.}\tabularnewline

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25cm² used to assess the presence or absence of 5 key classes of
vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm
over the sediment during low tide), Phaeophyceae (brown algae),
Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae
(red algae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded on the online on the Global Biodiversity Information Facility
(GBIF) platform \citep{BedeGbif}, a public open portal to store and
share biodiversity data. Each photograph was also processed to estimate
the percent cover of each type of vegetation using an image processing
software (imageJ). For all quadrat, the hyperspectral reflectance
signatures of each vegetation class was measured using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. The methodological scheme
followed in this study is presented in Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-vegetation}Class of vegetation used to train the
Neural Network model and their standardised spectral signature}

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights were utilized for training and
prediction of the Neural Network model, whereas low-resolution flights
were solely employed for prediction purposes.}

@@

The spectral similarities of the reflectance signatures between
intertidal green macrophytes (Magnoliopsida and Chlorophyceae) make
their discrimination challenging using simple classification algorithms
(Figure~\ref{fig-vegetation} F). To overcome this challenge, a deep
learning classification method was developed, trained, validated, and
applied to each drone flight (Figure~\ref{fig-workflow}).

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into 7 different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table 1) were used for training
purposes because their 8 mm spatial resolution allowied to avoid
spectral sub-pixel mixing and to accurately differentiate various
vegetation classes. More than 418,000 pixels at 8 mm resolution from the
3 training flights were used to train the model (Table 2). Twenty one
variables were used by the model as predictors: 10 raw spectral bands of
the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm
to 840 nm), the same 10 spectral bands standardized using a min/max
transformation (Equation~\ref{eq-std} ; \citep{Cao2017}) and the
Normalized difference vegetation index (NDVI, Equation~\ref{eq-ndvi}).
Standardisation of spectral bands is used to eliminate the scaling
differences between spectra and to limit the effect of biomass on the
shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

The model was applied to all flights at both 12 and 120 m of altitude.
In situ information on georeferenced class type and percent cover
collected during each flight was used to assess the model accuracy.
These images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset. A
confusion matrix, along with precision metrics such as global accuracy,
sensitivity, specificity, and Kappa coefficient, was generated for each
sites. All validation matrices were then merged to create a unique
matrix . Altogether, a total of 536,000 pixels was used to validate the
model, thus providing a geographically robust validation dataset.

@@

A total of 9 prediction maps corresponding to the 9 drone flights were
obtained. Each prediction map is associated with a probability map,
indicating the probability of the selected class for every pixel. The
low-altitude flight conducted in Gafanaha, Portugal, represents the site
with the highest complexity (Figure~\ref{fig-GafLow}). Among the 5
vegetation classes on which the model was trained, 4 were present on
this site. On this site, there is a mixture of Chlorophyceae and
Rhodophyceae over the seagrass meadow. This is also where
Bacillariophyceae is most abundant. Although the seagrass bed is solely
composed of \emph{Zostera noltei}, various colors can be observed: dark
green (indicating healthy beds with 100\% coverage) and whitish/brown
(indicating beds where leaves are dying). Regardless of the color, the
meadow is predicted as Magnoliopsida by the model.

@@

The high-altitude flight over Gafanha covered nearly 1 km² in total
(Figure~\ref{fig-GafHigh}). A channel delineating a small island was
masked in the prediction map due to sunglint and misclassification
caused by the turbid shallow water. Most of the intertidal area has been
classified as Magnoliopsida by the model, even though there is
discoloration of seagrass blades. Only a few pixels have been classified
as Chlorophyceae at this scale. Furthermore, the area that was
classified as Bacillariophyceae in the low-altitude flight remains
mostly classified as such in the high-altitude flight, though some
pixels were classified as Magnoliopsida. Patches of Rhodophyceae were
correctly classified. In the nothern part of the scene, near the land
limit, patches of \emph{Spartina sp.} were misclassified, either as
Magnoliopsida or as Phaeophyceae.

@@

The high altitude flight acquired over the inner lagoon of the Ria de
Aveiro is the largest of all flights, covering almost 1.5 km²
(Figure~\ref{fig-Boat}). On this site, only seagrass and red algae where
seen on the field. The classification provided consistent results, with
a patchy Magnoliopsida meadow mixed with Rhodophyceae on the eastern
part of the scene. As shown in the zoom (Figure~\ref{fig-Boat}), the
edges of the meadow can be colonised by Chlorophyceae.

@@

The flight over L'Epine in Noirmoutier Island, France
(Figure~\ref{fig-Dike}) was conducted near a dike crossing the northern
part of the scene from west to east. Alongside this dike, brown algae
attached to rocks and stranded green algae could be found. Despite the
high mixture between Chlorophyceae and Magnoliopsida these two classes
were correctly discriminated by the classifier.

@@

A total of 536,000 pixels was used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Kerdrean (5557 pixels) and Marinha Lanzarote (159713 pixels),
respectively. Model global accuracy was 94.26\% with a Kappa coefficient
of 0.92 (Figure~\ref{fig-Validation}). The least performing site was
Gafanha High (global accuracy of 75.45\%) whereas Mataducos was the site
with the most accurate prediction (global accuracy of 98.05\%). Overall,
the classes Phaeophyceae, Magnoliopsida, Sediment and Rhodophyceae were
correctly classified with a balanced accuracy of 1, 0.96, 0.96 and 0.91
respectively. Bacillariophyceae was the least performing class (accuracy
of 0.72 ) mainly due to the confusion between Magnoliopsida and
Sediment.

@@

The computation of the variable importance makes it possible to identify
which wavelengths are the most useful for class prediction
(Figure~\ref{fig-VIP}). The spectral bands at 444 , 717 and 842 nm of
the Micasense camera are important for none of the vegetation classes.
The band at 531 nm is the only important predictor for the classifier to
accurately predict chlorophyceae. In fact, at this wavelength, the
Chlorophyceae spectra has the maximum reflectance among of all
vegetation classes. The bands at 531 and 740 nm are the most important
predictors for Phaeophyceae, corresponding to the minimum reflectance
among all classes. Bands at 475 and 560 nm are the most important
predictors for Bacillariophyceae and Rhodophyceae, respectively. Four
predictors, ranging from the green (560 nm) to the RedEdge (705 nm)
bands are important to accurately predict magnoliopsida.

@@

Using the very high precision low altitude flight, we determined the
minimal percent cover required to classify correctly classify a given
class using the high altitude flight (Figure~\ref{fig-upscaling}). When
the percent cover of the class is 100 \%, big pixels are well classified
for all the classes excepted for Bare Sediment, where it's well
classified 80\% of the time. A vegetative percent cover of at least 80\%
is need to have all the big pixels well classified, at the exception of
Magnoliopsida that needs an higher percent cover (\textgreater90 \%) to
be well classified. Concerning the probability of each class, a really
high Percent cover is needed to confidently predict Bacillariophyceae.
To predict Chlorophyceae with a confidence of 0.85, a percent cover of
93 \% is needed, 90 \% for magnoliopsida, 92 \% for Rhodophyceae and 97
\% for Bacillariophyceae.

@@

The VIP analysis of the Neural Network model (see Fig. VIP) shows that
the 531 nm band is the most important spectral band for accurately
identifying Chlorophyceae. In fact, at this wavelength, Chlorophyceae
exhibits the highest reflectance among all other classes, highlighting
the difference in accessories carotenoid proportion between seagrasses
and green algae (\citep{repolho2017seagrass}).
